---
title: 'Statistical report: Temporal trend of heart failure hospitalizations in cardiology vs. non-cardiology wards in Sweden: data from the SwedeHF registry'
subtitle: 'DRAFT'
author: 'Statistician: Lina Benson'
  
date: "`r Sys.Date()`"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 7
    fig_width: 7
    number_sections: yes
link-citations: yes
bibliography: references.bib
nocite: '@*'
header-includes:
   - \usepackage{draftwatermark}
---

\newpage 
\tableofcontents 
\listoftables
\listoffigures
\newpage

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE, include = TRUE, comment = "",
  warning = FALSE, message = FALSE, fig.pos = "H",
  fig.path = "../output/figs/"
)
options(knitr.kable.NA = "")
```

```{r adjust_directory_if_needed, include=FALSE}
# Uncomment lines below if rmd file is placed in a subdirectory
knitr::opts_knit$set(root.dir = normalizePath("../"))
```

```{r load_project}
# 1. Set options in config/global.dcf
# 2. Load packages listed in config/global.dcf
# 3. Import functions and code in lib directory

ProjectTemplate::reload.project()

cacheon <- TRUE
```             

# Data handling

## Data source

Dataset prepared by Bodil Svennblad with corrections version 6 (rs.data6) with 
end follow-up 2016-12-31 consisting of data from SwedeHF, SCB (LISA Registry) 
and Socialstyrelsen (National Patient Registry and Cause of Death Registry). 
Note that ward type is no longer collected in SwedeHF due to problems with the 
definition and therefore data from the new SwedeHF is not included in the analysis. 
This also means that some care should be taken when interpreting the results. 

## Created variables 

Variables prepared in rs.data6 are documented elsewhere. This section specifies 
additional recoding and derivation of variables for the purpose of current 
analysis that are not considered to be self-explanatory. 

Income is considered below median if it is below the 50 percentile for that 
respective year of inclusion, and above median if it is above or equal to the 50 percentile. 

NT-proBNP is considered below median if it is below the 50 percentile 
and above median if it is above or equal to the 50 percentile within EF group. 

ras is defined as yes if arb, acei or arni = yes (and arb and acei both 
non-missing) and no if both arb and acei = no. Since arni was introduced 
into SwedeHF at a later date all missing are presumed to be no.   

Atrial_fib_flutter, IHD, Diabetes, Hypertension and valve disease are 
combined from registrations in SwedeHF and the derived variables from 
the National patient registry as follows: 

d_Atrial_fib_flutter is yes if any of the following are fulfilled: 
com_AF = "yes" or fibrillation = "yes" or ekg = "Atrial fibrillation",

d_IHD is yes if any of the following are fulfilled: 
com_IHD = "yes" or revascularized = "yes" or com_pci = "yes" or com_cabg = "yes 
(there is approx 70% missing in ihd in SwedeHF registations post 2013 so 
this variable was not included in the combinded variable),

d_Diabetes is yes if any of the following are fulfilled: 
com_DM = "yes" or diabetes = "yes", 

d_Hypertension is yes if any of the following are fulfilled: 
com_hypertension = "yes" or hypertension = "yes", 

d_Valvedisease is yes if any of the following are fulfilled: com_Valvular = "yes" 
or valveDisease = "yes" or heartValveSurgery = "yes"

If the variable in SwedeHF is missing (no missing in the National patient 
registry) the combined variable is also set to missing. 

## Inclusion/exclusion criteria

```{r flow}
default_kable(flow, caption = "Flowchart")
```

N unique patients in (). Observe that one patient can have multiple posts and 
therefore be counted more than once on each row (so for example can be in both the
excl cardiology cell and the excl medicine cell with different posts). 

First patient in: `r min(wdata$date)` and last patient in: `r max(wdata$date)`.  

The median age (IQR) is `r wdata %>% summarise(med = fn(median(age), dig = 1),
                                             q1 = fn(quantile(age, probs = 0.25), dig = 1),
                                             q3 = fn(quantile(age, probs = 0.75), dig = 1)) %>%
                                   mutate(out = paste0(med, " (", q1, "-", q3, ")")) %>%
                                   pull(out)` and 
`r wdata %>% count(gender) %>%
  mutate(perc = fn(n / sum(n) * 100, 1)) %>%
  filter(gender == "Female") %>%
  pull(perc)`% females.    

# Statistical analysis 

## General

All analyses were performed using `r sessionInfo()$R.version$version.string` [@r]. 
The level of significance is set to 5%, two-sided. No adjustment for multiple 
comparisons were made and therefore the results should be viewed with care.

## Time plot

```{r, child = "../src/barplot_cardio.Rmd"}

```

\clearpage
\newpage

## Baseline characteristics

```{r, child = "../src/tab1.Rmd"}

```

\clearpage
\newpage

# Reproducibility

## R session information {#sessioninfo}

```{r sessinfo}
sessionInfo()
```

## R code

The R code for all data handling and statistical analyses are found: 
https://github.com/KIHeartFailure/wardtypetemporaltrends. On publication
the repository will be made public so as to 
link to it from the resulting article for increased transparency and code sharing.
No data or output is stored in the repository. 

# References
